Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-03-03 ms.tex

Our analytic derviations neglect the effects of heat conduction, an assumption we now check.  The ratio of the magnitude of the conductive heating rate to the external heating rate at the stagnation radius is given by
\begin{align}
  \left.\frac{\nabla\cdot(\kappa \nabla T)}{q v_{\rm
w}^{2}/2}\right|_{r_{\rm s}} &\sim \frac{2t_{\rm h}\kappa_0
T|_{r_{\rm s}}^{7/2}}{r_{\rm s}^{2}\eta \rho_{\star}|_{r_{\rm s}} \tilde{v}_{w}^{2}(1+\phi)}
\nonumber \\ &\sim {\rm min}
  \begin{cases}
  10 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.8} v_{500}^{6.8} &  \text{unsaturated (core)}\\
 10 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.5} v_{500}^{5.4} &  \text{unsaturated (cusp)}\\
  1.0(\phi/0.1)/(2-\Gamma) & \text{saturated},
  \end{cases}
 \label{eq:conduction}
\end{align}
where the second equality makes use of equations (\ref{eq:Tanalytic}), (\ref{eq:stag_simple}), and we have approximated $\nabla^{2} \sim 1/r_{\rm s}^{2}$, $\nabla \sim 1/r_{\rm s}$.  We have approximated the stellar density profile as (eq.~[\ref{eq:rhostar}])
\begin{eqnarray}
  \rho_{\star}|_{r_{\rm s}} &\simeq& \frac{M_{\bullet}(2-\Gamma)}{4\pi r_{\rm soi}^{3}}\left(\frac{r_{\rm s}}{r_{\rm soi}}\right)^{-1-\Gamma} \nonumber \\
 &\approx& \begin{cases}
    6.1 \times 10^{-19}M_{\bullet,8}^{-1.2}v_{500}^{2.2}\,{\rm g\,cm^{-3}}
    & \text{core} \\
    2.1 \times 10^{-18} M_{\bullet,8}^{-1.5}v_{500}^{3.6}
    \,{\rm g\,cm^{-3}}  & \text{cusp}, 
  \end{cases}
  \label{eq:rhostarrs}
\end{eqnarray}
where the stagnation radius is assume to reside well inside the Nuker break radius.  Equation (\ref{eq:conduction}) shows that even in the case of saturated conduction, the neglect of heat conduction results in at most an order unity correction for expected values of the saturation parameter $\phi < 0.1$, justifying its neglect in our analysis.  Numerical experiments including conductivity confirm this ($\S\ref{sec:numerical}$).

@@

Radiative cooling can, however, be stabilized if it is overwhelmed by other sources of cooling, namely the term $\propto -q c_{s}^{2} \propto -qT$ in the entropy equation (eq.~[\ref{eq:dsdt}]).  Neglecting radiative cooling, this term is balanced at the stagnation radius by the wind heating term, $\dot{q}_{\rm heat} = q \tilde{v}_{w}^{2}/2$.  Thermal stability near the stagnation radius can thus be assessed by comparing the ratio of wind heating to radiative cooling $|\dot{q}_{\rm rad}|$ (eq.~[\ref{eq:qdot_rad}]),
\begin{align}
\left.\frac{\dot{q}_{\rm heat}}{|\dot{q}_{\rm rad}|}\right|_{\rm r_{\rm s}} \simeq
  \begin{cases}
   590 \eta_{0.02}^{-1} M_{\bullet,8}^{-0.76}v_{500}^{7.2}  &, \text{core}\\
   210 \eta_{0.02}^{-1} M_{\bullet,8}^{-0.48}v_{500}^{5.8}  &, \text{cusp},     
  \end{cases}
  \label{eq:cooling2}
\end{align}
where we have used equations (\ref{eq:rhors}) and (\ref{eq:rhostarrs}) for the gas and stellar densities at the stagnation radius, respectively.  We have approximated the cooling function for $T < 2\times 10^{7}$ K as $\Lambda(T) = 1.1 \times 10^{-22} \left(T/10^6 \text{K}\right)^{-0.7}  $erg cm$^3 $s$^{-1}$ assuming solar metallicity gas (\citealt{Draine:2011a}; his Fig. 34.1).

@@

The existence of a minimum heating rate for thermal stability (eq.~[\ref{eq:cooling2}]) implies a corresponding maximum accretion rate,
\begin{align}
\frac{\dot{M}_{\rm TI}}{\dot{M}_{\rm edd}} \simeq
  {\rm min}\begin{cases}
   4\times 10^{-4} \eta_{0.02}^{0.47} M_{\bullet,8}^{0.34} &, \text{core}\\
   2\times 10^{-4} \eta_{0.02}^{0.59} M_{\bullet,8}^{0.29}    &, \text{cusp}\\
\eta M_{\bullet}/\dot{M}_{\rm edd}t_{\rm h} \approx 6. 6\times 10^{-5}\eta_{0.02}&, v_{\rm TI} < \sigma,
  \end{cases}
  \label{eq:Mdotmax}
\end{align}
which is derived using equation (\ref{eq:eddr_analytic}) to calculate $\dot{M}(v_{\rm TI})$.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{rs.eps}
  \caption{\label{fig:stag} \emph{Top panel:} Stagnation radius
    $r_{s}$ in units of the sphere of influence radius $r_{\rm soi}$
    (eq.~[\ref{eq:rsoi}]) for galaxies in our sample as a function of
    the ratio of the effective wind heating rate $v_{w}$ to the
    stellar velocity dispersion $\sigma$.  Green, orange, and blue
    symbols correspond to different values of $v_{w} =$ 300, 600, and
    1200 km s$^{-1}$, respectively.  Squares correspond to cusp
    galaxies ($\Gamma = 0.8$), while triangles correspond to cores
    ($\Gamma = 0.1$). The black curves correspond to the analytic
    prediction from equation (\ref{eq:rs2main}), with solid and dashed
    curves calculated for parameters $\{\Gamma=0.8, \densSlope=1\}$ and
    $\{\Gamma=0.1,\densSlope=0.5\}$, respectively. }
\end{figure}

@@

\begin{eqnarray}
  \left.\frac{r_{\rm Ia}}{r_{\rm s}}\right|_{\rm v_{w}^{\rm Ia}}
  &\approx& 12\eta_{0.02}^{-1} M_{\bullet,8}^{-1.1}(\epsilon_{\rm
    Ia}/0.4) (\tau_{\star}/t_{\rm h})^{-0.56} \nonumber \\
  &\approx& 12 M_{\bullet,8}^{-1.1}(\epsilon_{\rm Ia}/0.4) (\tau_{\star}/t_{\rm h})^{0.74}
\label{eq:rIars}
\end{eqnarray}
If $r_{\rm Ia} \gg r_{\rm s}$, then Ia heating can only be
approximated as a steady heating source near the stagnation radius for
extremely massive SMBHs with $M_{\bullet} \gtrsim 10^9 \Msun$ or in
the case of a very young stellar population ($\tau_{\star} \ll t_{\rm
h}$).  On the other hand, equation (\ref{eq:rIars}) may underestimate
the importance of Ia heating because $r_{\rm Ia}/r_{\rm s}$ will be
substantially smaller than its value estimated using $v_{w}^{\rm Ia}$
if the heating rate excluding SN Ia is less than this.

@@

There are two types of SMBH feedback: kinetic and radiative.  Radiative feedback is potentially effective even in low luminosity AGN via Compton heating (e.g., \citealt{Sazonov+04}, \citealt{Ciotti+10}), which provides a volumetric heating rate (\citealt{Gan+14})
\be
\dot{e} = 4.1\times 10^{-35}n^{2}\xi T_{\rm C}\,{\rm erg\,cm^{-3}\,s^{-1}},
\ee
where $\xi = L/n r^{2}$ is the ionization parameter and $L$ is the SMBH luminosity with Compton temperature $T_{\rm C} \sim 10^{9}$ K $\gg T$ (e.g., \citealt{Ho99}, \citealt{Eracleous+10}).

The importance of Compton heating can be estimated by assuming the SMBH
radiates with a luminosity $L = \epsilon_{\rm rad} \dot{M}c^{2}$, where
$\epsilon_{\rm rad}$ is the radiative efficiency and where $\dot{M}$ is
estimated from equation (\ref{eq:mdot_analytic}).  Then using
equations (\ref{eq:rs_simple}), (\ref{eq:mdot_analytic}),
(\ref{eq:rhors}), (\ref{eq:rhostarrs}) we calculate that from equation
(\ref{eq:vw_eff}) that the effective heating rate at the stagnation radius is given by
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 29 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.38}v_{500}^{-1.4}\,{\rm
km\,s^{-1}} &, \text{core}\\ 50 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.24}v_{500}^{-0.7}\,{\rm
km\,s^{-1}} &, \text{cusp},
  \end{cases}
  \label{eq:vC}
\end{align} where $T_{C,9} = T_{C}/10^{9}$ K and $\epsilon_{-2} =
\epsilon_{\rm rad}/0.01 \sim 1$ is the maximum radiative efficiency
for an advective accretion flow expected to produce high $T_{\rm C}$
emission (e.g.~\citealt{Narayan&Yi95}).  We caveat that, unlike
stellar wind heating, Compton heating depends on radius, scaling as
$v_{w}^{\rm C}(r) \propto (n^{2}\xi/\rho_{\star})^{1/2} \propto
r^{(\Gamma-\densSlope-1)/2}$, i.e. $\propto r^{-0.6}$ and $\propto r^{-0.7}$
for core ($\Gamma = 0.8$; $\densSlope = 1$) and cusp ($\Gamma = 0.1$;
$\densSlope = 0.5$) galaxies, respectively.  The assumption of our model that the heating
parameter be constant with radius is not satisfied, but this radial
variation is sufficiently weak that it should not significantly alter
our conclusions.

If Compton heating acts alone, i.e. $\tilde{v}_{w} = v_{w}^{C}$, then solving equation
(\ref{eq:vC}) for $\tilde{v}_{\rm w}$ yields
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 150 \eta_{0.02}^{0.21}T_{\rm
C,9}^{0.21}\epsilon_{-2}^{0.21} M_{\bullet,8}^{0.16}\,{\rm km\,s^{-1}}
&, \text{core}\\ 130 \eta_{0.02}^{0.29}T_{\rm
C,9}^{0.29}\epsilon_{-2}^{0.29} M_{\bullet,8}^{0.14}\,{\rm km\,s^{-1}}
&, \text{cusp},
  \end{cases}
  \label{eq:vC2}
\end{align} 
Compton heating can thus be significant in young stellar populations with relatively high mass loss rates, e.g. $v_{w}^{\rm C} \gtrsim 300$ km s$^{-1}$ for $\eta \gtrsim 1$.  The mass accretion rate corresponding to the state where Compton heating limits the average accretion rate is given by substituting equation (\ref{eq:vC2}) into equation (\ref{eq:eddr_analytic}):
\begin{align}
\frac{\dot{M}_{C}}{\dot{M}_{\rm edd}} \approx 
\begin{cases} 1.3\times 10^{-3} \eta_{0.02}^{0.20}T_{\rm
C,9}^{-0.8}\epsilon_{-2}^{-0.8} M_{\bullet,8}^{0.15}
&, \text{core}\\ 6\times 10^{-4} \eta_{0.02}^{0.30}T_{\rm
C,9}^{-0.7}\epsilon_{-2}^{-0.7} M_{\bullet,8}^{0.14}
&, \text{cusp},
  \end{cases}
  \label{eq:MdotCa}
\end{align}
If the accretion luminosity $\epsilon = \dot{M}/\dot{M}_{\rm edd}$ depends on the Eddington ratio as predicted by some models for radiatively inefficient accretion flows (\citealt{Narayan&Yi95}; \citealt{Narayan+98}), then the Compton-limited accretion rate becomes
\begin{align}
\frac{\dot{M}_{C}}{\dot{M}_{\rm edd}} \approx 
\begin{cases} 3\times 10^{-3} \eta_{0.02}^{0.11}T_{\rm
C,9}^{-0.44}M_{\bullet,8}^{0.08}
&, \text{core}\\ 2\times 10^{-3} \eta_{0.02}^{0.18}T_{\rm
C,9}^{-0.4} M_{\bullet,8}^{0.08}
&, \text{cusp}.
  \end{cases}
  \label{eq:MdotC}
\end{align}

>>>>

Our analytic derviations neglect the effects of heat conduction, an assumption we now check.  The ratio of the magnitude of the conductive heating rate to the external heating rate at the stagnation radius is given by
\begin{align}
  \left.\frac{\nabla\cdot(\kappa \nabla T)}{q v_{\rm
w}^{2}/2}\right|_{r_{\rm s}} &\sim \frac{2t_{\rm h}\kappa_0
T|_{r_{\rm s}}^{7/2}}{r_{\rm s}^{2}\eta \rho_{\star}|_{r_{\rm s}}
\tilde{v}_{w}^{2}}
 \left(\frac{1}{1+\frac{\kappa_0 T^{7/2}|_{r_{\rm s}}}{5 \rs
    \phi \rho c_s^3}}\right)
\nonumber \\ &\sim {\rm min}
  \begin{cases}
  20 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.8} v_{500}^{6.8} &  \text{unsaturated (core)}\\
 20 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.5} v_{500}^{5.4} &  \text{unsaturated (cusp)}\\
  2 \phi & \text{saturated},
  \end{cases}
 \label{eq:conduction}
\end{align}
where the second equality makes use of equations (\ref{eq:Tanalytic}),
(\ref{eq:stag_simple}), and we have approximated $\nabla^{2} \sim
1/r_{\rm s}^{2}$, $\nabla \sim 1/r_{\rm s}$.  We have approximated the
stellar density profile as (eq.~[\ref{eq:rhostar}])
\begin{eqnarray}
  \rho_{\star}|_{r_{\rm s}} &\simeq& \frac{M_{\bullet}(2-\Gamma)}{4\pi r_{\rm soi}^{3}}\left(\frac{r_{\rm s}}{r_{\rm soi}}\right)^{-1-\Gamma} \nonumber \\
 &\approx& \begin{cases}
    6.1 \times 10^{-19}M_{\bullet,8}^{-1.2}v_{500}^{2.2}\,{\rm g\,cm^{-3}}
    & \text{core} \\
    2.1 \times 10^{-18} M_{\bullet,8}^{-1.5}v_{500}^{3.6}
    \,{\rm g\,cm^{-3}}  & \text{cusp}, 
  \end{cases}
  \label{eq:rhostarrs}
\end{eqnarray}
where the stagnation radius is assume to reside well inside the Nuker
break radius.  Equation (\ref{eq:conduction}) shows that even in the
case of saturated conduction, the neglect of heat conduction results
in at most an order unity correction for expected values of the
saturation parameter $\phi < 0.1$, justifying its neglect in our
analysis.  Numerical experiments including conductivity confirm this
($\S\ref{sec:numerical}$). {\bf AG: If we make this statement then we
  should be sure to include the numerical experiments. The original
  formula was not correct for the saturated case. The numerical values
also changed somewhat. I would appreciate it if someone could double
check this. Also note that there is a weak core/cusp dependence in the
saturated case which I have chosen not to include.}

@@

Radiative cooling can, however, be stabilized if it is overwhelmed by other sources of cooling, namely the term $\propto -q c_{s}^{2} \propto -qT$ in the entropy equation (eq.~[\ref{eq:dsdt}]).  Neglecting radiative cooling, this term is balanced at the stagnation radius by the wind heating term, $\dot{q}_{\rm heat} = q \tilde{v}_{w}^{2}/2$.  Thermal stability near the stagnation radius can thus be assessed by comparing the ratio of wind heating to radiative cooling $|\dot{q}_{\rm rad}|$ (eq.~[\ref{eq:qdot_rad}]),
\begin{align}
\left.\frac{\dot{q}_{\rm heat}}{|\dot{q}_{\rm rad}|}\right|_{\rm r_{\rm s}} \simeq
  \begin{cases}
   430 \eta_{0.02}^{-1} M_{\bullet,8}^{-0.76}v_{500}^{7.2}  &, \text{core}\\
   840 \eta_{0.02}^{-1} M_{\bullet,8}^{-0.48}v_{500}^{5.8}  &, \text{cusp},     
  \end{cases}
  \label{eq:cooling2}
\end{align}
where we have used equations (\ref{eq:rhors}) and (\ref{eq:rhostarrs})
for the gas and stellar densities at the stagnation radius,
respectively.  We have approximated the cooling function for $T <
2\times 10^{7}$ K as $\Lambda(T) = 1.1 \times 10^{-22} \left(T/10^6
  \text{K}\right)^{-0.7} $erg cm$^3 $s$^{-1}$ assuming solar
metallicity gas (\citealt{Draine:2011a}; his Fig. 34.1).

@@

The existence of a minimum heating rate for thermal stability (eq.~[\ref{eq:cooling2}]) implies a corresponding maximum accretion rate,
\begin{align}
\frac{\dot{M}_{\rm TI}}{\dot{M}_{\rm edd}} \simeq
  {\rm min}\begin{cases}
   6\times 10^{-4} \eta_{0.02}^{0.47} M_{\bullet,8}^{0.34} &, \text{core}\\
   2\times 10^{-4} \eta_{0.02}^{0.59} M_{\bullet,8}^{0.29}    &, \text{cusp}\\
\eta M_{\bullet}/\dot{M}_{\rm edd}t_{\rm h} \approx 6. 6\times 10^{-5}\eta_{0.02}&, v_{\rm TI} < \sigma,
  \end{cases}
  \label{eq:Mdotmax}
\end{align}
which is derived using equation (\ref{eq:eddr_analytic}) to calculate $\dot{M}(v_{\rm TI})$.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{rs.eps}
  \caption{\label{fig:stag} \emph{Top panel:} Stagnation radius
    $r_{s}$ in units of the sphere of influence radius $r_{\rm soi}$
    (eq.~[\ref{eq:rsoi}]) for galaxies in our sample as a function of
    the ratio of the effective wind heating rate $v_{w}$ to the
    stellar velocity dispersion $\sigma$.  Green, orange, and blue
    symbols correspond to different values of $v_{w} =$ 300, 600, and
    1200 km s$^{-1}$, respectively.  Squares correspond to cusp
    galaxies ($\Gamma = 0.8$), while triangles correspond to cores
    ($\Gamma = 0.1$). The black curves correspond to the analytic
    prediction from equation (\ref{eq:rs2main}), with solid and dashed
    curves calculated for parameters $\{\Gamma=0.8, \densSlope\simeq1\}$ and
    $\{\Gamma=0.1,\densSlope\simeq0.6\}$, respectively. }
\end{figure}

@@

\begin{align}
  \left.\frac{r_{\rm Ia}}{r_{\rm s}}\right|_{\rm v_{w}^{\rm Ia}}&\approx
  \begin{cases}
    8\, \eta_{0.02}^{-1} M_{\bullet,8}^{-1.1}(\epsilon_{\rm
     Ia}/0.4) (\tau_{\star}/t_{\rm h})^{-0.56}  & {\rm core}\\
    17\, \eta_{0.02}^{-1} M_{\bullet,8}^{-1.1}(\epsilon_{\rm
     Ia}/0.4) (\tau_{\star}/t_{\rm h})^{-0.56} &{\rm cusp}\\
   \end{cases} \nonumber\\
 &\approx 
 \begin{cases}
    8\, \eta_{0.02}^{-1} M_{\bullet,8}^{-1.1}(\epsilon_{\rm
     Ia}/0.4) (\tau_{\star}/t_{\rm h})^{0.74}  & {\rm core}\\
    17\, \eta_{0.02}^{-1} M_{\bullet,8}^{-1.1}(\epsilon_{\rm
     Ia}/0.4) (\tau_{\star}/t_{\rm h})^{0.74} &{\rm cusp}\\
   \end{cases}
\label{eq:rIars}
\end{align}
If $r_{\rm Ia} \gg r_{\rm s}$, then Ia heating can only be
approximated as a steady heating source near the stagnation radius for
extremely massive SMBHs with $M_{\bullet} \gtrsim 10^9 \Msun$ or in
the case of a very young stellar population ($\tau_{\star} \ll t_{\rm
h}$).  On the other hand, equation (\ref{eq:rIars}) may underestimate
the importance of Ia heating because $r_{\rm Ia}/r_{\rm s}$ will be
substantially smaller than its value estimated using $v_{w}^{\rm Ia}$
if the heating rate excluding SN Ia is less than this.

@@

There are two types of SMBH feedback: kinetic and radiative.  Radiative feedback is potentially effective even in low luminosity AGN via Compton heating (e.g., \citealt{Sazonov+04}, \citealt{Ciotti+10}), which provides a volumetric heating rate (\citealt{Gan+14})
\begin{align}
\dot{e} = 4.1\times 10^{-35}n^{2}\xi T_{\rm C}\,{\rm erg\,cm^{-3}\,s^{-1}},
\end{align}
where $\xi = L/n r^{2}$ is the ionization parameter and $L$ is the SMBH luminosity with Compton temperature $T_{\rm C} \sim 10^{9}$ K $\gg T$ (e.g., \citealt{Ho99}, \citealt{Eracleous+10}).

The importance of Compton heating can be estimated by assuming the SMBH
radiates with a luminosity $L = \epsilon_{\rm rad} \dot{M}c^{2}$, where
$\epsilon_{\rm rad}$ is the radiative efficiency and where $\dot{M}$ is
estimated from equation (\ref{eq:mdot_analytic}).  Then using
equations (\ref{eq:rs_simple}), (\ref{eq:mdot_analytic}),
(\ref{eq:rhors}), (\ref{eq:rhostarrs}) we calculate that from equation
(\ref{eq:vw_eff}) that the effective heating rate at the stagnation radius is given by
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 27 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.38}v_{500}^{-1.4}\,{\rm
km\,s^{-1}} &, \text{core}\\ 30 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.24}v_{500}^{-0.7}\,{\rm
km\,s^{-1}} &, \text{cusp},
  \end{cases}
  \label{eq:vC}
\end{align} where $T_{C,9} = T_{C}/10^{9}$ K and $\epsilon_{-2} =
\epsilon_{\rm rad}/0.01 \sim 1$ is the maximum radiative efficiency
for an advective accretion flow expected to produce high $T_{\rm C}$
emission (e.g.~\citealt{Narayan&Yi95}).  We caveat that, unlike
stellar wind heating, Compton heating depends on radius, scaling as
$v_{w}^{\rm C}(r) \propto (n^{2}\xi/\rho_{\star})^{1/2} \propto
r^{(\Gamma-\densSlope-1)/2}$, i.e. $\propto r^{-0.6}$ and $\propto r^{-0.75}$
for core ($\Gamma = 0.8$; $\densSlope \simeq 1$) and cusp ($\Gamma = 0.1$;
$\densSlope \simeq 0.6$) galaxies, respectively.  The assumption of our model that the heating
parameter be constant with radius is not satisfied, but this radial
variation is sufficiently weak that it should not significantly alter
our conclusions.

If Compton heating acts alone, i.e. $\tilde{v}_{w} = v_{w}^{C}$, then solving equation
(\ref{eq:vC}) for $\tilde{v}_{\rm w}$ yields
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 150 \eta_{0.02}^{0.21}T_{\rm
C,9}^{0.21}\epsilon_{-2}^{0.21} M_{\bullet,8}^{0.16}\,{\rm km\,s^{-1}}
&, \text{core}\\ 96 \eta_{0.02}^{0.29}T_{\rm
C,9}^{0.29}\epsilon_{-2}^{0.29} M_{\bullet,8}^{0.14}\,{\rm km\,s^{-1}}
&, \text{cusp},
  \end{cases}
  \label{eq:vC2}
\end{align} 
Compton heating can thus be significant in young stellar populations with relatively high mass loss rates, e.g. $v_{w}^{\rm C} \gtrsim 300$ km s$^{-1}$ for $\eta \gtrsim 1$.  The mass accretion rate corresponding to the state where Compton heating limits the average accretion rate is given by substituting equation (\ref{eq:vC2}) into equation (\ref{eq:eddr_analytic}):
\begin{align}
\frac{\dot{M}_{C}}{\dot{M}_{\rm edd}} \approx 
\begin{cases} 3\times 10^{-3} \eta_{0.02}^{0.20}T_{\rm
C,9}^{-0.8}\epsilon_{-2}^{-0.8} M_{\bullet,8}^{0.15}
&, \text{core}\\ 8\times 10^{-4} \eta_{0.02}^{0.30}T_{\rm
C,9}^{-0.7}\epsilon_{-2}^{-0.7} M_{\bullet,8}^{0.14}
&, \text{cusp},
  \end{cases}
  \label{eq:MdotCa}
\end{align}
If the accretion luminosity $\epsilon = 0.1 \dot{M}/\dot{M}_{\rm edd}$ depends on the Eddington ratio as predicted by some models for radiatively inefficient accretion flows (\citealt{Narayan&Yi95}; \citealt{Narayan+98}), then the Compton-limited accretion rate becomes
\begin{align}
\frac{\dot{M}_{C}}{\dot{M}_{\rm edd}} \approx 
\begin{cases} 1.4\times 10^{-2} \eta_{0.02}^{0.11}T_{\rm
C,9}^{-0.44}M_{\bullet,8}^{0.08}
&, \text{core}\\ 6\times 10^{-3} \eta_{0.02}^{0.18}T_{\rm
C,9}^{-0.4} M_{\bullet,8}^{0.08}
&, \text{cusp}.
  \end{cases}
  \label{eq:MdotC}
\end{align}
{\bf AG: Note that I have included the bolometric correction for now.}

### 2015-03-03 windheat.tex
more in 5

For a stellar population that forms impulsively, the rate of mass and energy injection per unit stellar mass at age $t$ are given, respectively, by 
\begin{align} 
  \dot{\bar{m}}(t) &= \frac{\Delta M(t) \mu|_{M_{\rm TO}(t)}
    \left|\dot{M}_{\rm TO}(t)\right| + f_{\rm MS} \int_{m_0}^{m_{\rm
        T}(t)}
    \dot{m}(\Mstar, t) \mu|_{\Mstar} {\rm d}\Mstar }{\bar{m}_*}\\
  \dot{\bar{e}}(t) &=\dot{e}_{\rm TO}(t)+ f_{\rm MS} \int_{m_0}^{m_{\rm T}(t)}
  \frac{\vw^2(\Mstar, t) \dot{m}(\Mstar, t) \mu|_{\Mstar} {\rm d}\Mstar}{\bar{m}_*},
  \label{eq:edotImp}
\end{align} 
where the first and second terms in each expression corresponds to
contributions from main sequence (MS) stars and post-main sequence
(PMS) stars, respectively.  Here $ f_{\rm MS} < 1$ is the efficiency
with which main sequence winds thermalize their energy with the rest
of the injected gas (we hereafter assume $f_{\rm MS} = 1$) and $\mu$
is the IMF, which we take to be of the Salpeter form $\mu|_{\Mstar} \propto
M_{\star}^{-2.35}$, truncated at $m_0 = 0.1 \Msun$ on the low mass end and at
$100 \Msun$ on the high mass end, where $\bar{m}_*$ = 0.35 $\Msun$ is
the corresponding mean stellar mass.

@@

\subsection{Over Star Formation History}
Generalizing to an arbitrary star formation history $S(t)$, the total rate of mass and energy input can be written as
\begin{align} 
  \dot{M}(t) &= \int_0^t S(t_1) \dot{\bar{m}}(t-t_1){\rm
      d}t_1\\
  \dot{E}(t) &= \int_0^t S(t_1) \dot{\bar{e}}(t-t_1){\rm
      d}t_1,\\
\end{align}
resulting in a wind heating parameter of 
\begin{align}
  v_w^2(t) &=2 \dot{E}(t)/\dot{M}(t).
\end{align}

The mass return parameter will be

\begin{align}
\eta = \frac{\dot{M}(t)}{\int_0^t S(t_1) {\rm d}t_1} \th
\end{align}

% \begin{figure}
% \includegraphics[width=\columnwidth]{vw.pdf}
% \caption{\label{NickPlot2} Effective wind velocities $V_{\rm w}$ for
% different $S(t)$.  The yellow and orange curves are the Moster SF
% histories, with and without Type II SNe, respectively.  The red,
% purple, and blue curves are Moster SFs convolved with a $\sin^2(t)$
% function normalized to $10^7$, $10^8$, and $10^9$ yr fluctuations,
% respectively.  These solid curves lack SN, but the dashed purple curve
% possesses it.  Effective $V_{\rm w}$ is strongly diminished when the
% variability timescale is greater than $\sim$ twice the duration of
% high-velocity winds.}
% \end{figure}

We estimate the stellar wind heating provided by the {\it average}
star formation history of galaxies of a given $\Mbh$ using the results
of \citet{MosterNaab+:2013a} (their eqs.~[17]-[20]).  Note that the
star formation histories in \citet{MosterNaab+:2013a} are in terms of
halo mass. For a given $\Mbh$ we assign the halo mass whose star
formation history would produce a bulge consistent with the
$\Mbh-M_{\rm bulge}$ relationship from \citet{McConnellMa:2013a}.

>>>>

For a stellar population that forms impulsively, the rate of mass and energy injection per unit stellar mass at age $t$ are given, respectively, by 
\begin{align} 
  \dot{\bar{m}}(t) &= \frac{\Delta M(t) \mu|_{M_{\rm TO}(t)}
    \left|\dot{M}_{\rm TO}(t)\right| + f_{\rm MS} \int_{m_0}^{m_{\rm
        T}(t)}
    \dot{m}(\Mstar, t) \mu|_{\Mstar} {\rm d}\Mstar }{\bar{m}_*}\\
  \dot{\bar{e}}(t) &=\dot{e}_{\rm TO}(t)+ f_{\rm MS} \int_{m_0}^{m_{\rm T}(t)}
  \frac{\vw^2(\Mstar, t) \dot{m}(\Mstar, t) \mu|_{\Mstar} {\rm d}\Mstar}{\bar{m}_*},
  \label{eq:edotImp}
\end{align} 
where the first and second terms in each expression corresponds to
contributions from main sequence (MS) stars and post-main sequence
(PMS) stars, respectively.  Here $ f_{\rm MS} < 1$ is the efficiency
with which main sequence winds thermalize their energy with the rest
of the injected gas (we hereafter assume $f_{\rm MS} = 1$) and $\mu$
is the IMF, which we take to be of the Salpeter form $\mu|_{\Mstar} \propto
M_{\star}^{-2.35}$, truncated at $m_0 = 0.1 \Msun$ on the low mass end and at
$100 \Msun$ on the high mass end, where $\bar{m}_*$ = 0.35 $\Msun$ is
the corresponding mean stellar mass.  Our optimistic assumption of 100 per cent thermalization efficiency of stellar winds is at least partially offset by our conservative assumption to neglect energy input from core collapse supernovae, which contribute a comparable energy injection rate to massive stellar winds.

@@

\subsection{Over Star Formation History}
Generalizing to an arbitrary star formation history $S(t)$, the total rate of mass and energy input can be written as
\begin{align} 
  \dot{M}(t) &= \int_0^t S(t_1) \dot{\bar{m}}(t-t_1){\rm
      d}t_1 \label{eq:MDotSFH}\\
  \dot{E}(t) &= \int_0^t S(t_1) \dot{\bar{e}}(t-t_1){\rm
      d}t_1, \label{eq:EDotSFH}
\end{align}
resulting in a wind heating parameter of 
\begin{align}
  v_w^2(t) &=2 \dot{E}(t)/\dot{M}(t).
\end{align}
The mass return parameter will be
\begin{align}
\eta = \frac{\dot{M}(t)}{\int_0^t S(t_1) {\rm d}t_1} \th
\end{align}
 % \begin{figure}
 % \includegraphics[width=\columnwidth]{NickPlot2.pdf}
 % \caption{\label{NickPlot2} Effective wind velocities for
 % nonstandard star formation histories.  The black curve shows, for reference, $v_{\rm w}$ calculated using the halo-averaged $S(t)$.  Other colors show the effective wind velocities $\mathcal{V}_{\rm w}$ computed using convolved star formation histories $\mathcal{S}(t)$.  The purple curve shows a recent drop in the star formation rate, i.e. $t_{\rm scale}=10^6~{\rm yr}$.  The blue and red curves show $t_{\rm scale}=10^7~{\rm yr}$ and $t_{\rm scale}=10^8~{\rm yr}$, respectively; longer scale times look quite similar to the red curves.  Dashed, dotted, and dot-dashed line styles show star formation rates that have declined to fractions $\epsilon = 0.001$, $\epsilon = 0.1$, and $\epsilon = 0.3$ of their halo-averaged values, respectively. }
 % \end{figure}
We estimate the stellar wind heating provided by the {\it average}
star formation history of galaxies of a given $\Mbh$ using the results
of \citet[eqs.~17-20]{MosterNaab+:2013a}.  Note that the
star formation histories in \citet{MosterNaab+:2013a} are in terms of
halo mass. For a given $\Mbh$ we assign the halo mass whose star
formation history would produce a bulge consistent with the
$\Mbh-M_{\rm bulge}$ relationship from \citet{McConnellMa:2013a}.  A slight complication occurs for the largest mass halos, where much of the $z=0$ stellar mass has been acquired through accretion of satellite halos rather than {\it in situ} star formation.  To accomodate this, we use analytic fits for mass accretion histories $A(t)$, also taken from \citet[eqs.~21-23]{MosterNaab+:2013a}, and replace $S(t) \rightarrow S(t)+A(t)$ in Eqs. \ref{eq:MDotSFH} and \ref{eq:EDotSFH}.

Figure \ref{NickPlot2} shows how the wind heating varies as star formation histories deviate from their halo-averaged values.  In particular, we show
\begin{align} 
  \dot{\mathcal{M}}(t) &= \int_0^t \mathcal{S}(t_1) \dot{\bar{m}}(t-t_1){\rm
      d}t_1\\
  \dot{\mathcal{E}}(t) &= \int_0^t \mathcal{S}(t_1) \dot{\bar{e}}(t-t_1){\rm
      d}t_1,\\
  \mathcal{V}_w^2(t) &=2 \dot{\mathcal{E}}(t)/\dot{\mathcal{M}}(t).
\end{align}
In these equations, 
\begin{equation}
\mathcal{S}(t) = S(t) \times \left(\frac{2}{\pi}(1-\epsilon)\arctan(t/t_{\rm scale}) + \epsilon \right).
\end{equation}
This function convolves the recent ($z \approx 0$) halo-averaged star formation history with local variation to give a more pessimistic estimate for the value of $\tilde{v}_{\rm w}$.  In particular, replacing $S(t)$ with $\mathcal{S}(t)$ reduces the recent star formation rate to a fraction $\epsilon$ of its halo-averaged value, and does so for a characteristic time $t_{\rm scale}$ into the past.  As we can see in Fig. \ref{NickPlot2}, this dramatically lowers the effective wind speed when both $t_{\rm scale} \gtrsim 10^7 ~{\rm yr}$ and $\epsilon \lesssim 0.1$, but otherwise has too modest of an effect to change the thermal stability properties of the flow (although the location of $r_{\rm s}$ and the value of $\dot{M}$ may change significantly).
